%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Answers for 6.863 Assignment 6
% Created by Salman Ahmad
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[10pt]{article}

\setlength{\topmargin}{-0.8in}
\setlength{\oddsidemargin}{0in}
\setlength{\textwidth}{6.5in}
\setlength{\textheight}{9.25in}
\setlength{\parindent}{0in}

\usepackage{sectsty}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{subfigure}
\usepackage{longtable}

\def\arraystretch{1.5}%  1 is the default, change whatever you need


\usepackage{array,etoolbox}
\preto\tabular{\setcounter{magicrownumbers}{0}}
\newcounter{magicrownumbers}
\newcommand\rownumber{\stepcounter{magicrownumbers}\arabic{magicrownumbers}}



\sectionfont{\large}
\subsectionfont{\normalsize}
\subsubsectionfont{\small}

\newcommand{\handout}[3]{
  \renewcommand{\thepage}{#1 - \arabic{page}}
  \noindent
  \begin{center}
    \hspace*{-0.25in}\framebox[6.5in]{
      \vbox{
        \hbox to 6.25in { {\bf 6.863J/9.611J: Natural Language Processing } 
                          \hfill Prof. Robert C. Berwick }
        \vspace{4mm}
        \hbox to 6.25in { {\Large \hfill #1  \hfill} }
        \vspace{2mm}
        \hbox to 6.25in { {\it #2 \hfill #3} }
        }
      }
  \end{center}
  \vspace*{4mm}
}

\begin{document}

\handout{Assignment 6: Statistical Parsing and Lexical Semantics}{My posse consists of: {\tt null}}{Salman Ahmad (saahmad@mit.edu)}

\section*{Part II: Semantic Analysis of ``{\tt Invest}''}

\subsection*{Data Collection}

To collect all of the occurrences of ``invest'' along with their subcategorization frame I used the follow {\bf {\tt tregex}} query on the supplied subset of the Penn Tree Bank:

{\tt
\begin{verbatim}
@VP < ( __< /^[Ii]nvest[esi]$|^[Ii]nvest$/ )
\end{verbatim}
}

The query returned 52 matches in 52 trees.

\subsection*{Analysis}

A summary of the data is shown in the next section.

\hspace{1in}

According to Levin the word ``invest'' is mainly used as an ``fulfilling'' verb that is used to describe something that will be provided or utilized in the future. ``Invest'' can be used in two subtle different contexts: it can be used to specify the {\bf act} that is being provided and it can also be used to specify the {\bf thing or item} that is being provided. The verb typically appears in a ``NP V NP to / with NP'' frame. For example:


\begin{itemize}
\item ``John invested the land to Sally''
\end{itemize}


Looking at the subcategories in the Data Tabulation section you can see that not all of the frames follow the ``NP V NP to / with NP'' subcategory that Levin describes. In fact, many of them do not even include a prepositional phrase at all, for example, subcategories (5, 7, 18) and many others. These usages are not described in Levin at all. Examples of uses of ``invest'' that are not accompanied by a preposition include:

\begin{itemize}
\item ``invest internationally''
\item ``invest elsewhere''
\end{itemize}

Furthermore, Levin talks at length about the use of the ``to'' and ``with'' prepositions that are used with verbs like ``Invest''. However, this almost never appears in the Penn Tree Bank. Rather, the Penn Tree Bank is dominated by the use of the ``in'' preposition with invest. There are plenty of examples:

\begin{itemize}
\item ``invest money in stocks that they 'll need in the near future -- for example , for college tuition payments or retirement expenses''
\item ``invest in common stocks''
\item ``invest in business''
\end{itemize}

In summary, Levin primarily describes ``invest'' being used as an ``equip'' and ``fulfilling'' verb while in practice it seems rarely used in the context.

\subsection*{Data Tabulation}

The table below displays the number of times a unique subcategorization frame exists in the Penn Tree Bank. 

\begin{centering}


\begin{longtable}{ l l l }

\caption{Subcategory counts}\\
\hline\hline
\textbf{Subcategory} & \textbf{Count} \\
\hline\hline
\endfirsthead
\multicolumn{2}{c}
{\tablename\ \thetable\ -- \textit{Continued from previous page}} \\
\hline\hline
\textbf{Subcategory} & \textbf{Count} \\
\hline\hline
\endhead
\hline \multicolumn{2}{r}{\textit{Continued on next page}} \\
\endfoot
\hline\hline
\endlastfoot
\hline

\rownumber)\space\space\space\space [VP (VB invest) PP-CLR] & 12 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) NP PP-CLR] & 8 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) ADVP-LOC] & 2 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) NP PP-LOC-CLR] & 2 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) ] & 2 \\ \hline
\rownumber)\space\space\space\space [VP (VBP invest) PP-CLR] & 2 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) , SBAR-ADV] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) ADVP-MNR PP-CLR PP-TMP] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) ADVP-MNR] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) NP ADVP-LOC PP-MNR] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) NP ADVP-LOC-CLR] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) NP PP-CLR , NP-2] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) NP PP-CLR , PP-2] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) NP PP-CLR S-PRP] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) NP PP-CLR SBAR-1] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) NP PP-TMP PP-LOC-CLR S-CLR] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) NP S-PRP] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) NP] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) PP-CLR PP-PRP] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) PP-LOC-CLR NP] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) PP-LOC-CLR] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VB invest) PP-TMP] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VBP invest) ADVP PP-LOC-CLR ADVP-LOC] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VBP invest) ADVP-LOC] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VBP invest) PP-LOC-CLR] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VBZ invests) ADVP PP-CLR] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VBZ invests) ADVP PP-LOC-CLR] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VBZ invests) ADVP PP-LOC] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VBZ invests) NP NP-TMP PP-CLR SBAR-ADV] & 1 \\ \hline
\rownumber)\space\space\space\space [VP (VBZ invests) NP PP-CLR] & 1 \\ \hline

\end{longtable}
\end{centering}


\end{document}

